---
title: "Raster test"
output: html_notebook
---

```{r housekeeping}
library(tidyverse)
library(raster)
library(rgdal)
library(here)
library(reshape)
library(scales)
library(R.utils)
library(sp)
library(sf)
library(parallel)
library(lubridate)
library(lfstat)
library(stringr)
```

Master notebook for raster transformations, Rhone project. See raster_test.rmd for applying Data Carpentry workshop to Rhone data, see DC_raster_tutorial.rmd for finished workshop modules

Projection used is Lambert Conformal Conic (code 32610)

Mtg w/ Ryan 8/7/20:

Make a bounding box (could be general and make it generous so reproject will remain insite) or specific coordinates
-bbox function can get bounding box
-project polyline/shapefile into bbox

*eeflux - precip = water use

Look at correlations, look @ chirps lit on FTP server
-dopar library (parallel processing R), splits for-loop (slow) into groups of tasks, distributed to processer cores
-can raster stack the chirps data & do analysis that way, parallelized correlation on 2 time series

```{r reproject raster}
# Data Carpentry reproject raster tutorial - Rhone data https://datacarpentry.org/r-raster-vector-geospatial/03-raster-reproject-in-r/index.html

eeflux_1985 <- raster(here("rasters/pdr_eeflux/1985_07_08_et.tif"))

chirps_1985 <- raster(x = here("rasters/chirps/chirps-v2.0.2017.04.16.tif"))
                      #xmn = 4,
                      #xmx = 6,
                      #ymn = 43,
                      #ymx = 47) # Lat/long bounding box for all 3 of Bryn's sites - doesn't change anything

chirps_1985_utm <- projectRaster(from = chirps_1985, # This seems to work to snap chirps to eeflux
                                 to = eeflux_1985,
                                 alignOnly = TRUE)
# Specifying individual attributes from eeflux - not necessary?
                                #crs = crs(eeflux_1985),
                                #extent = extent(eeflux_1985)
                                #res = res(eeflux_1985))

eeflux_1985_df <- as.data.frame(eeflux_1985, xy = TRUE) %>% 
  mutate(mmh2o = X1985_07_08_et)

chirps_1985_df <- as.data.frame(chirps_1985_utm, xy = TRUE) %>% 
  mutate(mmh2o = chirps.v2.0.2017.04.16)

# Checking coordinates, extent
crs(eeflux_1985)

rhone_res <- res(eeflux_1985)

crs(chirps_1985_utm)

res(chirps_1985_utm)

# Another command to get crs/projection: proj4string(eeflux_1985)

extent(eeflux_1985)

extent(chirps_1985_utm)

# Checking to see if reprojection worked:
ggplot() +
  geom_raster(data = chirps_1985_df,
              aes(x = x,
                  y = y,
                  fill = mmh2o)) +
  coord_quickmap() #+
  theme(axis.title = element_blank()) +
  scale_fill_manual(values = terrain.colors(4),
                    name = "Precipitation")
```


```{r CHIRPS bbox test (not used except for coordinate info)}
# Initial bbox approach + notes from Casey
# Trying to create a bounding box according to this approach: https://twitter.com/TimSalabim3/status/1063099774977667072?s=20
# (bbox function is no longer supported)
# Don't think this works w/ raster data

tif <- raster(here("rasters/pdr_eeflux/1985_07_08_et.tif"))
chirps_crs <- crs(chirps)
#make extent here - chirps in wgs84?
# keep in mind reprojection method - nearest neighbor?
#raster::extent(xmin, xmax, ymin, ymax)

# None of this is necessary \/
new_bb = c(4, 6, 43, 47)
names(new_bb) = c("xmin", "ymin", "xmax", "ymax")
attr(new_bb, "class") = "bbox"

attr(st_geometry(tif),
  "bbox") = new_bb

st_bbox(tif)

# Clipping extent by filtering chirps df by degrees:

chirps_test <- raster(x = here("rasters/chirps/1990/chirps-v2.0.1990.07.16.tif"))

chirps_test_df <- as.data.frame(chirps_test, xy = TRUE)

chirps_test_tbl <- as_tibble(chirps_test_df) %>% 
  filter(chirps.v2.0.1990.07.16 != "-9999") %>% 
  filter(x >= 4) %>% 
  filter(x <= 7) %>% 
  filter(y >= 43) %>% 
  filter(y <= 47)

ggplot() +
  geom_raster(data = chirps_test_tbl,
              aes(fill = chirps.v2.0.1990.07.16,
                  x = x,
                  y = y)) +
  coord_quickmap() +
  theme(axis.title = element_blank())

# Getting bounding box from PDR (eeflux tifs from multiple download sessions had same bounding box)

# Entire river is in UTM zone 31 - reprojecting eeflux from utm to lambert (rhone valleys) makes them line up

test_85 <- raster(here("rasters/pdr_eeflux/2019_08_23_et.tif"))

st_bbox(test_85)

bbox3 <- st_as_sfc(st_bbox(test_85))

sf::st_write(bbox3, here("rasters/pdr_eeflux/pdr_bbox/bbox_pdr.shp"), driver = "ESRI Shapefile")

pdr.forest <- st_read(here("rasters/bbox/pdr_bbox/bd_foret_pdr.shp"))

# Lambert conformal conic projection w/ 2 standard parallels @ 49degN/44degN
rhone.crs <- as.character(crs(pdr.forest))

test_85 <- projectRaster(from = test_85,
                         crs = rhone_crs)

test_85_df <- as.data.frame(test_85, xy = TRUE)

chirps_test_lambert <- projectRaster(from = chirps_test, crs = rhone_crs, res = res(test_85))

chirps_test_df <- as.data.frame(chirps_test_lambert, xy = TRUE)

# This is producing no values here? Works above?
chirps_test_tbl <- as_tibble(chirps_test_df) %>% 
  filter(chirps.v2.0.1990.07.16 != "-9999") %>% 
  filter(x >= 4) %>% 
  filter(x <= 7) %>% 
  filter(y >= 43) %>% 
  filter(y <= 47)

ggplot() + 
  geom_raster(data = chirps_test_df,
              aes(fill = chirps.v2.0.1990.07.16,
                  x = x,
                  y = y)) +
  geom_raster(data = test_85_df,
              aes(x = x,
                  y = y,
                  fill = X2019_08_23_et)) +
  geom_sf(data = pdr_forest) +
  coord_sf() 
```

``` {r gz unzip (for CHIRPS data)}
# Unzipping .gz files (download format for chirps) - finished
  
chirps <- list.files(path = here("rasters/chirps"), 
                     pattern = ".gz", 
                     full.names = TRUE, 
                     recursive = TRUE)

for(file in chirps){
  gunzip(file)
}

list.dirs(path = "serdp/")

```
 
```{r clipping/reprojecting test}
# Chirps calculations - clipping/reprojecting function

# From Ryan phone call - find Lambert conical crs, try to clip/reproject/re-resolution brick of chirps

chirps_1981 <- list.files(path = here("rasters/chirps/1981"), 
                     pattern = ".tif", 
                     full.names = TRUE)

NAvalue(x) <- -9999.0


  #z <- projectRaster(from = x,
                # res = rhone_res, - want to do this after finding averages?
                # crs = rhone_crs,
                #overwrite = TRUE)

chirps_rmna_save <- function(x) {
  writeRaster(x, overwrite = TRUE)
}
  
  projectRaster()

lapply_test <- chirps_clip(here("rasters/chirps/chirps-v2.0.1990.07.16.tif"))
  #chirps_clip(here("/rasters/chirps/1981/chirps-v2.0.1981.01.02.tif"))

# chirps_test <- raster(here("rasters/chirps-clipped/chirps-v2.0.2020.06.30-clipped.tif"))

# eeflux_test <- raster(here("rasters/eeflux/BRC/2011-09-28-ET-BRC.tif")) %>% 
  projectRaster(crs = crs(chirps_test))

crs(eeflux_test) <- crs()

lapply_test_df <- as.data.frame(lapply_test, xy = TRUE)

res(lapply_test)

crs(lapply_test)

lapply_test

# Test works
ggplot() +
  geom_raster(data = lapply_test_df,
              aes(x = x,
                  y = y,
                  fill = mmh2o)) +
  coord_quickmap()

raster(here("rasters/chirps/1981/chirps-v2.0.1981.01.01.tif"))

parallel::mclapply(x = chirps_1981,
                   )

chirps_test_stack <- stack(chirps_1981)

# think you can brick a list w/o making stack first (stack takes way less time - can do transformations on a stack?)

chirps_test_brick <- brick(chirps_test_stack) # Hung on anvil for almost 2hrs

parallel::mclapply()

```
 
```{r final CHIRPS clipping fxn from global to watershed scale}
# Final chirps clipping function (w/ help from Casey O'Hara)
rhone.bb <- matrix(c(4,43,6,47), nrow = 2)
rhone.area <- raster::extent(rhone.bb)

# Making same bbox for geojson to download new ET data
x_coord <- c(4, 6)
y_coord <- c(43, 47)
xym <- cbind(x_coord, y_coord)
xym

p = sp::Polygon(xym)
ps = Polygons(list(p),1)
sps = SpatialPolygons(list(ps))
plot(sps)

# Creating blank bbox using OGR 

data = data.frame(f=99.9)
spdf = SpatialPolygonsDataFrame(sps,data)
rhone.area2 <- spdf
summary(spdf)
extent(spdf)

writeOGR(rhone.area2, here("rasters/bbox/rhonebb.GeoJSON"),
         layer = 'rhone.area2',
         driver = "GeoJSON", 
         overwrite = TRUE
      )

chirps.clip.SA <- 
  # This function clips global chirps rasters to the Rhone watershed study area
  function(x) {
    
  outfile.base <- str_replace(basename(x), '.tif', '-clipped.tif')
  outfile <- here('rasters/chirps-clipped', outfile.base)
  
  if(file.exists(outfile)) {
    message('File already exists! Skipping..', outfile)
  }
  else{message('Processing file: ', outfile)
    rast <- raster(x) %>% 
      crop(x, y = rhone.area, overwrite = TRUE) %>% 
      setNames('mmh2o')
    
    values(rast)[values(rast) < 0] <- NA
    
    rast.out <- raster::projectRaster(from = rast,
                              crs = rhone.crs,
                              method = 'ngb')
    
    writeRaster(rast.out, 
                filename = outfile,
                overwrite = TRUE)
  }
  return(NULL)
  }

chirps <- list.files(path = here("rasters/chirps"), 
                     pattern = ".tif", 
                     full.names = TRUE, 
                     recursive = TRUE)
 
mclapply(chirps, chirps.clip.SA) # Use multi-core apply function on server for speed
```
 
```{r unzip/rename EEFLUX}
# Writing a function to rename eeflux

# Make folder for each site, dump .zip folders, create list of files

ISP_list <- list.files(here("rasters/eeflux/ISP"), 
                       pattern = ".zip",
                       full.names = TRUE)

auto_unzip <- function(x){for (zip in x)
  x <- unzip(zip, overwrite = TRUE, exdir = here("rasters/eeflux/ISP"))
  # Use lubridate - how to specify character positions? Also rename PDR rasters? For loop?
}

# Write rename function, code help from Marc
auto_rename <- function(x){for (tif in x)
  year <- substr(x, start=63, stop=66)
  day <- substr(x, start = 67, stop = 69) %>% 
    as.numeric()
  nye <- "01-01"
  date <- as.Date(day, origin = paste(year, nye, sep = "-"))
  
  site <- "BRC"
  
  name <- paste0(date, "-ET-", site)
  path <- paste0(dirname(x), "/", name, ".tif")

  file.rename(x, path)
}

ISP_tif <- list.files(here("rasters/eeflux/ISP"), 
                       pattern = ".tif",
                       full.names = TRUE)

BRC_tif <- list.files(here("rasters/eeflux/BRC"),
                      pattern = ".tif",
                      full.names = TRUE)

```
 
```{r crop to forested valley areas (mask imported from Arc) + precip aggregation}
# Function to stack chirps by year, crop to pdr valleys

pdr.valley.shp <- sf::st_read(here("rasters/bbox/pdr_bbox/bd_foret_pdr.shp")) #%>% 
  st_transform(crs = crs(eeflux_temp))
# Use sf package, readOGR gives weird file structure, sf should work as crop()
# Convert shp from LCC to UTM (eeflux) - don't have to reproject chirps (but clipped rasters are already in lcc!)
  
isp.valley.shp <- sf::st_read(here("rasters/bbox/isp_bbox/bd_foret_isp_filtered.shp"))

brc.valley.shp <- sf::st_read(here("rasters/bbox/brc_bbox/bd_foret_brc_filtered.shp"))

pdr.valley.shp <- sf::st_read(here("rasters/bbox/pdr_bbox/bd_foret_pdr_filtered.shp"))

# OGR functions = workaround for when sf not working, not recommended

isp.valley.shp2 <- readOGR( 
  dsn= here("rasters/bbox/isp_bbox/bd_foret_isp_filtered.shp"), 
  #layer="bd_foret_isp",
  verbose=FALSE
)

brc.valley.shp2 <- readOGR( 
  dsn= here("rasters/bbox/brc_bbox/bd_foret_brc_filtered.shp"), 
  #layer="bd_foret_brc",
  verbose=FALSE
) # From https://www.r-graph-gallery.com/168-load-a-shape-file-into-r.html

# Set to updated valley mask
eeflux_temp <- raster(here('rasters/eeflux/ISP/2020-08-27-ET-ISP.tif')) %>% 
  projectRaster(crs = crs(isp.valley.shp),
                extent = extent(isp.valley.shp)) %>% 
  raster::mask(mask = isp.valley.shp)

# Create second mask to run simultaneously on anvil
eeflux_temp_brc <- raster(here('rasters/eeflux/BRC/2020-07-26-ET-BRC.tif')) %>% 
  projectRaster(crs = crs(brc.valley.shp),
                extent = extent(brc.valley.shp)) %>% 
  raster::mask(mask = brc.valley.shp)

# PDR mask
eeflux_temp_pdr <- raster(here('rasters/eeflux/PDR/2007_03_16_et.tif')) %>% 
  projectRaster(crs = crs(pdr.valley.shp),
                extent = extent(pdr.valley.shp)) %>% 
  raster::mask(mask = pdr.valley.shp)

eeflux_temp_brc2 <- raster(here('rasters/bbox/brc_bbox/brc_crop.tif'))

writeRaster(eeflux_temp_brc, here("rasters/bbox/brc_bbox/brc_crop2.tif"), format = "GTiff")
  
chirps.clipped <- list.files(here('rasters/chirps-clipped'),
                             full.names = TRUE)

crs(raster(here('rasters/chirps-clipped/chirps-v2.0.2020.06.30-clipped.tif')))

# This command returns raster_list df w/ file path + water year
raster_df <- data.frame(f = chirps.clipped) %>% # Take chirps.clipped, give name f
  mutate(wateryear = lfstat::water_year(x =
    lubridate::ymd(
      str_extract(chirps.clipped, pattern = "[0-9]{4}\\.[0-9]{2}\\.[0-9]{2}")),
    origin = 'usgs'))

class(raster_df) # Want to feed raster a character vector (filenames)

wy1981 <- raster_df %>% 
  filter(wateryear == 1981) %>% 
  dplyr::select(-wateryear) %>% 
  .$f #. says take object piped in and give column f
  
x = wy1981$f

water_years <- c(1981:2020)

# This function calculates total precip per water year - runs very slowly (4hrs per year)

crop_stack <- function(y){
  
  for (y in water_years){
    
    message('Year is now ', y)
  
    wy <- raster_df %>% 
      filter(wateryear == y) %>%
      dplyr::select(-wateryear) %>% 
      .$f %>% 
      as.vector()
    
  suppressWarnings({ # This says run wystack, and don't give crs warnings (will run faster)
      wystack <- raster::stack(wy)
    })
  
    wystack_2 <- projectRaster(wystack, to = eeflux_temp_brc2) # converted eeflux to lcc proj above
    
    wystack_3 <- mask(wystack_2, mask = brc.valley.shp2) # change for new site
    
    wy_total <- wystack_3 %>% calc(fun = sum, na.rm = TRUE)
    
    # Change for new site
    writeRaster(wy_total,
                filename = here('rasters/chirps-wy/chirps-brc', 
                                paste("brc_chirps_", y, ".tif", sep = '')),
                overwrite = TRUE)
  
  }
return(plot(wystack_2))
}
    
read_test <- raster(here('rasters/chirps-wy/chirps-pdr/pdr_chirps_1981.tif'))

extent_test <- projectRaster(raster(here('rasters/chirps-clipped/chirps-v2.0.2020.06.30-clipped.tif')),
                             to = eeflux_temp)

plot(wy_total)

    #outfile <- here('rasters/chirps-wy/chirps-pdr', 'chirps-pdr-wy'y)
  ##outfile.base <- chr(y)
  ##outfile <- here('rasters/chirps-wy/chirps-pdr',outfile.base)

stack_test <- raster::stack(wy1981)
res(stack_test) <- 30
crop(wy-stack, y = pdr.valley.shp, overwrite = TRUE)
  #wy-total <- wy-stack %>% 
    #calc(year, fun = sum, na.rm = TRUE)

mclapply(water_years,
         crop_stack,
         mc.silent = FALSE)

crop_stack(1981)

```

```{r raster math raw calculations - not integrated}
plot(raster(here("rasters/chirps-wy/chirps-pdr/pdr_chirps_2019.tif")))

# Raster math

eeflux_pdr_2019 <- stack(
  list.files(
  here("rasters/eeflux/PDR"),
  pattern = "2019.",
  full.names = TRUE)) 
  
eeflux_pdr_2019 <- projectRaster(from = eeflux_pdr_2019,
              to = chirps_pdr_2019) %>% 
  calc(fun = sum, na.rm = TRUE)

chirps_pdr_2019 <- raster(here("rasters/chirps-wy/chirps-pdr/pdr_chirps_2019.tif"))

crs(eeflux_pdr_2019)

crs(chirps_pdr_2019)

correlate = (eeflux_pdr_2019
             #*(365/8)) 
             - chirps_pdr_2019)

correlate_masked <- mask(correlate, mask = pdr.valley.shp)

correlate_df <- as.data.frame(correlate,
                              xy = TRUE)

ggplot() +
  geom_raster(data = correlate_df,
              aes(x = x,
                  y = y,
                  fill = layer)) +
    coord_quickmap()
```

```{r import centerline from Arc, transect test - not used}
#Adding centerline to raster

centerline <- sf::read_sf(here("rasters/Rhone_centerline/Rhone_centerline.shp")) %>% 
  st_sf() %>% 
  st_transform(crs = crs(correlate)) %>% 
  as_Spatial() %>% 
  as('SpatialLines')

#not working
centerline_proj <- as.character(
  sf_project(from = centerline,
             to = correlate))

#not working
raster::distance(x = correlate,
                 y = centerline)

#Approach from https://gis.stackexchange.com/questions/226554/create-raster-in-r-distance-to-line-feature
require(rgeos)
dd = gDistance(centerline, as(correlate,"SpatialPoints"), byid=TRUE)
# This creates a matrix with a column for each feature in Sl. To get the nearest distance to any feature, apply min over rows:

correlate[] = apply(dd,1,min)
plot(correlate)

dist_masked <- mask(correlate, mask = pdr.valley.shp)

# More raster math
dist_corr = dist_masked / correlate_masked
plot(dist_corr)
```

```{r raster math with integrated annual site rasters from Python}
# Growing season calculations - https://statnmap.com/2018-01-27-spatial-correlation-between-rasters/
brc.valley.utm <- st_transform(brc.valley.shp, 32610)
isp.valley.utm <- st_transform(isp.valley.shp, 32610)

eeflux_75_rasters <- list.files(path = here("rasters/ee_growing_season_integrated_pdr"), 
                     pattern = ".tif", 
                     full.names = TRUE)

eeflux_75_stack <- stack(eeflux_75_rasters) %>% 
  projectRaster(to = eeflux_temp_pdr)

eeflux_75_mask <- mask(eeflux_75_stack, mask = pdr.valley.shp)

chirps_75_rasters <- list.files(path = here("rasters/ee_season_precip_data_pdr"), 
                     pattern = ".tif", 
                     full.names = TRUE)

chirps_75_stack <- stack(chirps_75_rasters) %>% 
  projectRaster(to = eeflux_temp_pdr)

chirps_75_mask <- mask(chirps_75_stack, mask = pdr.valley.shp)

#Pixel-wise regression, from: https://matinbrandt.wordpress.com/2014/05/26/pixel-wise-regression-between-two-raster-time-series/
s <- stack(eeflux_75_mask, chirps_75_mask)

fun=function(x) { if (is.na(x[1])){ NA } else { lm(x[1:35] ~ x[36:70])$coefficients[2] }} 
# Important to check that correct rasters are referenced in lm ^ check years in stack using s[1]
slope <- calc(s, fun)
writeRaster(x = slope,
              filename = here("rasters/raster_download/slope_125_275_pdr.tif"),
              format = 'GTiff',
            overwrite = TRUE)

fun=function(x) { if (is.na(x[1])){ NA } else { lm(x[1:35] ~ x[36:70])$coefficients[1] }}
intercept <- calc(s, fun)

fun=function(x) { if (is.na(x[1])){ NA } else { m <- lm(x[1:35] ~ x[36:70]);summary(m)$r.squared }}
r.squared <- calc(s, fun)
writeRaster(x = r.squared,
              filename = here("rasters/raster_download/r_squared_125_275_pdr.tif"),
              format = 'GTiff',
            overwrite = TRUE)

# Integrating p-values into existing workflow using: https://stackoverflow.com/questions/5587676/pull-out-p-values-and-r-squared-from-a-linear-regression
fun=function(x) { if (is.na(x[1])){ NA } else { m <- lm(x[1:35] ~ x[36:70]);
f <- summary(m)$fstatistic
	p <- pf(f[1],f[2],f[3],lower.tail=F)
	attributes(p) <- NULL
	return(p) }}
p.val <- calc(s, fun)
writeRaster(x = p.val,
              filename = here("rasters/raster_download/p_val_125_275_pdr.tif"),
              format = 'GTiff',
            overwrite = TRUE)

#Try each seasonal regression - April/May/June
chirps_75_spring <- list.files(path = here("4q_seasonal_data"), 
                     pattern = "amj.tif", 
                     full.names = TRUE) %>% 
  stack() %>% 
  projectRaster(to = eeflux_temp) %>% 
  mask(mask = pdr.valley.utm)

s_amj <- stack(eeflux_75_mask, chirps_75_spring)

fun=function(x) { if (is.na(x[1])){ NA } else { m <- lm(x[1:35] ~ x[36:70]);summary(m)$r.squared }}
r.squared_amj <- calc(s_amj, fun)

#R-squared is lower for AMJ than 125-275 comparison

chirps_75_winter <- list.files(path = here("4q_seasonal_data"), 
                     pattern = "jfm.tif", 
                     full.names = TRUE) %>% 
  stack() %>% 
  projectRaster(to = eeflux_temp) %>% 
  mask(mask = pdr.valley.utm)

s_jfm <- stack(eeflux_75_mask, chirps_75_winter)

fun=function(x) { if (is.na(x[1])){ NA } else { m <- lm(x[1:35] ~ x[36:70]);summary(m)$r.squared }}
r.squared_jfm <- calc(s_jfm, fun)

# Even lower for JFM

chirps_75_summer <- list.files(path = here("4q_seasonal_data"), 
                     pattern = "jas.tif", 
                     full.names = TRUE) %>% 
  stack() %>% 
  projectRaster(to = eeflux_temp) %>% 
  mask(mask = pdr.valley.utm)

s_jas <- stack(eeflux_75_mask, chirps_75_summer)

fun=function(x) { if (is.na(x[1])){ NA } else { m <- lm(x[1:35] ~ x[36:70]);summary(m)$r.squared }}
r.squared_jas <- calc(s_jas, fun)

chirps_75_fall <- list.files(path = here("4q_seasonal_data"), 
                     pattern = "ond.tif", 
                     full.names = TRUE) %>% 
  stack() %>% 
  projectRaster(to = eeflux_temp) %>% 
  mask(mask = pdr.valley.utm)

s_ond <- stack(eeflux_75_mask, chirps_75_fall)

fun=function(x) { if (is.na(x[1])){ NA } else { m <- lm(x[1:35] ~ x[36:70]);summary(m)$r.squared }}
r.squared_ond <- calc(s_ond, fun)

# All 4 seasonal correlations have lower R-squared than annual-to-annual
```

