---
title: "Raster test"
output: html_notebook
---

```{r}
#Housekeeping
library(tidyverse)
library(raster)
library(rgdal)
library(here)
library(reshape)
library(scales)
library(R.utils)
library(sp)
library(sf)
library(parallel)
library(lubridate)
```

Master notebook for raster transformations, Rhone project. See raster_test.rmd for applying Data Carpentry workshop to Rhone data, see DC_raster_tutorial.rmd for finished workshop modules

Mtg w/ Ryan 8/7:

Make a bounding box (could be general and make it generous so reproject will remain insite) or specific coordinates
-bbox function can get bounding box
-project polyline/shapefile into bbox

*eeflux - precip = water use

Look at correlations, look @ chirps lit on FTP server
-dopar library (parallel processing R), splits for-loop (slow) into groups of tasks, distributed to processer cores
-can raster stack the chirps data & do analysis that way, parallelized correlation on 2 time series

```{r}
# Data Carpentry reproject raster tutorial - Rhone data https://datacarpentry.org/r-raster-vector-geospatial/03-raster-reproject-in-r/index.html

eeflux_1985 <- raster(here("rasters/pdr_eeflux/1985_07_08_et.tif"))

chirps_1985 <- raster(x = here("rasters/chirps/chirps-v2.0.2017.04.16.tif"))
                      #xmn = 4,
                      #xmx = 6,
                      #ymn = 43,
                      #ymx = 47) # Lat/long bounding box for all 3 of Bryn's sites - doesn't change anything

chirps_1985_utm <- projectRaster(from = chirps_1985, # This seems to work to snap chirps to eeflux
                                 to = eeflux_1985,
                                 alignOnly = TRUE)
# Specifying individual attributes from eeflux - not necessary?
                                #crs = crs(eeflux_1985),
                                #extent = extent(eeflux_1985)
                                #res = res(eeflux_1985))

eeflux_1985_df <- as.data.frame(eeflux_1985, xy = TRUE) %>% 
  mutate(mmh2o = X1985_07_08_et)

chirps_1985_df <- as.data.frame(chirps_1985_utm, xy = TRUE) %>% 
  mutate(mmh2o = chirps.v2.0.2017.04.16)

# Checking coordinates, extent
crs(eeflux_1985)

rhone_res <- res(eeflux_1985)

crs(chirps_1985_utm)

res(chirps_1985_utm)

# Another command to get crs/projection: proj4string(eeflux_1985)

extent(eeflux_1985)

extent(chirps_1985_utm)

# Checking to see if reprojection worked:
ggplot() +
  geom_raster(data = chirps_1985_df,
              aes(x = x,
                  y = y,
                  fill = mmh2o)) +
  coord_quickmap() #+
  theme(axis.title = element_blank()) +
  scale_fill_manual(values = terrain.colors(4),
                    name = "Precipitation")
```


```{r}
# Initial bbox approach + notes from Casey
# Trying to create a bounding box according to this approach: https://twitter.com/TimSalabim3/status/1063099774977667072?s=20
# (bbox function is no longer supported)
# Don't think this works w/ raster data

tif <- raster(here("rasters/pdr_eeflux/1985_07_08_et.tif"))
chirps_crs <- crs(chirps)
#make extent here - chirps in wgs84?
# keep in mind reprojection method - nearest neighbor?
#raster::extent(xmin, xmax, ymin, ymax)

# None of this is necessary \/
new_bb = c(4, 6, 43, 47)
names(new_bb) = c("xmin", "ymin", "xmax", "ymax")
attr(new_bb, "class") = "bbox"

attr(st_geometry(tif),
  "bbox") = new_bb

st_bbox(tif)

# Clipping extent by filtering chirps df by degrees:

chirps_test <- raster(x = here("rasters/chirps/1990/chirps-v2.0.1990.07.16.tif"))

chirps_test_df <- as.data.frame(chirps_test, xy = TRUE)

chirps_test_tbl <- as_tibble(chirps_test_df) %>% 
  filter(chirps.v2.0.1990.07.16 != "-9999") %>% 
  filter(x >= 4) %>% 
  filter(x <= 7) %>% 
  filter(y >= 43) %>% 
  filter(y <= 47)

ggplot() +
  geom_raster(data = chirps_test_tbl,
              aes(fill = chirps.v2.0.1990.07.16,
                  x = x,
                  y = y)) +
  coord_quickmap() +
  theme(axis.title = element_blank())

# Getting bounding box from PDR (eeflux tifs from multiple download sessions had same bounding box)

# Entire river is in UTM zone 31 - reprojecting eeflux from utm to lambert (rhone valleys) makes them line up

test_85 <- raster(here("rasters/pdr_eeflux/2019_08_23_et.tif"))

st_bbox(test_85)

bbox3 <- st_as_sfc(st_bbox(test_85))

sf::st_write(bbox3, here("rasters/pdr_eeflux/pdr_bbox/bbox_pdr.shp"), driver = "ESRI Shapefile")

pdr_forest <- st_read(here("rasters/pdr_eeflux/pdr_bbox/bd_foret_pdr.shp"))

rhone_crs <- as.character(crs(pdr_forest))

test_85 <- projectRaster(from = test_85,
                         crs = rhone_crs)

test_85_df <- as.data.frame(test_85, xy = TRUE)

chirps_test_lambert <- projectRaster(from = chirps_test, crs = rhone_crs, res = res(test_85))

chirps_test_df <- as.data.frame(chirps_test_lambert, xy = TRUE)

# This is producing no values here? Works above?
chirps_test_tbl <- as_tibble(chirps_test_df) %>% 
  filter(chirps.v2.0.1990.07.16 != "-9999") %>% 
  filter(x >= 4) %>% 
  filter(x <= 7) %>% 
  filter(y >= 43) %>% 
  filter(y <= 47)

ggplot() + 
  geom_raster(data = chirps_test_df,
              aes(fill = chirps.v2.0.1990.07.16,
                  x = x,
                  y = y)) +
  geom_raster(data = test_85_df,
              aes(x = x,
                  y = y,
                  fill = X2019_08_23_et)) +
  geom_sf(data = pdr_forest) +
  coord_sf() 
```

``` {r}
# Unzipping .gz files - finished
  
chirps <- list.files(path = here("rasters/chirps"), 
                     pattern = ".gz", 
                     full.names = TRUE, 
                     recursive = TRUE)

for(file in chirps){
  gunzip(file)
}

list.dirs(path = "serdp/")

```
 
```{r}
# Chirps calculations - clipping/reprojecting function

# From Ryan phone call - find Lambert conical crs, try to clip/reproject/re-resolution brick of chirps

chirps_1981 <- list.files(path = here("rasters/chirps/1981"), 
                     pattern = ".tif", 
                     full.names = TRUE)


rhone_bb <- matrix(c(4,43,6,47), nrow = 2)
rhone_area <- raster::extent(rhone_bb)

NAvalue(x) <- -9999.0

chirps_clip <- function(x) {x <- raster(x)
  x <- crop(x, y = rhone_area)
  values(x)[values(x) < 0] <- NA
  names(x) <- "mmh2o"
  #z <- projectRaster(from = x,
                # res = rhone_res, - want to do this after finding averages?
                # crs = rhone_crs,
                #overwrite = TRUE)
  return(x)
} 

chirps_rmna_save <- function(x) {
  writeRaster(x, overwrite = TRUE)
}
  
  projectRaster()

lapply_test <- chirps_clip(here("rasters/chirps/chirps-v2.0.1990.07.16.tif"))
  #chirps_clip(here("/rasters/chirps/1981/chirps-v2.0.1981.01.02.tif"))

lapply_test_df <- as.data.frame(lapply_test, xy = TRUE)

res(lapply_test)

crs(lapply_test)

lapply_test

# Test works
ggplot() +
  geom_raster(data = lapply_test_df,
              aes(x = x,
                  y = y,
                  fill = mmh2o)) +
  coord_quickmap()

raster(here("rasters/chirps/1981/chirps-v2.0.1981.01.01.tif"))

parallel::mclapply(x = chirps_1981,
                   )

chirps_test_stack <- stack(chirps_1981)

# think you can brick a list w/o making stack first (stack takes way less time - can do transformations on a stack?)

chirps_test_brick <- brick(chirps_test_stack) # Hung on anvil for almost 2hrs

parallel::mclapply()

```
 
```{r}
# Writing a function to rename eeflux

# Make folder for each site, dump .zip folders, create list of files

auto_eeflux <- function(x) {for .zip in (x)
  y <- unzip(x)
  y <- raster(y)
  # Use lubridate - how to specify character positions? Also rename PDR rasters? For loop?
}

```
 
 